---
title: "p8105_hw1_jy3269"
author: "Jingyi Yao"
date: "`r Sys.Date()`"
output: github_document
---

#### Load packages 
```{r,results='hide',warning=FALSE,message=FALSE}
library(tidyverse)
```


## Problem 1

#### Load the `penguins` data set from the `palmerpenguins` package
```{r}
data("penguins", package = "palmerpenguins")   # load the penguins data set

penguins

```

#### Description of the `Penguins`data set

1. The `penguins` data set has `r ncol(penguins)` columns and `r nrow(penguins)` rows

2. The column names of the data set are 
  
    `r colnames(penguins)`

3. The mean flipper length is  `r mean(penguins$flipper_length_mm,na.rm =TRUE)`


#### Scatter Plot
```{r}
ggplot(penguins,aes(x = bill_length_mm,y = flipper_length_mm,colour = species)) + geom_point()+
  labs(title="flipper_length_mm (y) vs bill_length_mm (x)")
                                             # make a scatter plot

ggsave("scatter_plot_flipper_vs_bill.pdf")   # save the plot in pdf format

```


## Problem 2

#### Create a dataframe
```{r}
df_1 <- tibble(                  # use tibble() to create a dataframe named df_1
  norm_samp = rnorm(n = 10),     # sample from standard normal distribution
  logic_vec = norm_samp > 0,     # logical vector
  char_vec = letters[1:10],      # character vector of the first 10 English letters
  fac_vec = factor( c("male","female","transgender","male","male","transgender","female","female","female","female"))               # factor vector of 3 types of gender
)
 
df_1                             # show the created dataframe

```


#### Caculate the mean of variables _Before_ converting the types
```{r}
sample_mean <- mean(pull(df_1,var=1)) # mean of norm_samp in df_1
sample_mean

logic_mean <- mean(pull(df_1,var=2))   # mean of logic_vec in df_1
logic_mean

char_mean <- mean(pull(df_1,var=3))   # mean of char_vec in df_1
char_mean

fac_mean <- mean(pull(df_1,var=4))    # mean of fac_vec in df_1
fac_mean

```

#### Convert the types of variables
```{r,results='hide',warning=FALSE}
logic_vec_new <- as.numeric(pull(df_1,var=2))   # converted logical vector

char_vec_new <- as.numeric(pull(df_1,var=3))    # converted character vector

fac_vec_new <- as.numeric(pull(df_1,var=4))     # converted factor vector

```


#### Caculate the mean of variables _After_ converting the types
```{r}
print(logic_mean_new<-mean(logic_vec_new)) # mean of converted logical vector

print(char_mean_new<-mean(char_vec_new))   # mean of converted character vector

print(fac_mean_new<-mean(fac_vec_new))     # mean of converted factor vector

```

After the `as.numeric()` conversion to the variables:

1. `r df_1$logic_vec`    is turned into   `r logic_vec_new`,
    *  `TRUE` is given the value of `1` 
    *  `FALSE` is given the value of `0`

2. `r df_1$char_vec`   is turned into  `r char_vec_new`
    *  The warning shows that the characters are not transformed into numbers using `as.numeric()`
    *  We only get NAs in the new vector

3. `r df_1$fac_vec`   is turned into `r fac_vec_new`

   *  `female` is given the value of `1`
   *  `male` is given the value of `2`
   *  `transgender` is given the value of `3`





